Brief Announcement: Tight Bounds for Online Balanced Repartitioning

### 2020-05-21 podc_NEW/podc_ba.tex
BA 3 pages

We present the algorithm \PPL that is  asymptotically optimal for the learning problem.
It utilizes the concept of components,
introduced by Avin~et~al.~\cite{repartition-disc}.

\noindent
\textbf{Perfect Partition Learner algorithm.}
On each inter-cluster request $\{u,v\}$, 
\PPL  merges
 the two components that contain nodes $u$ and $v$.
To collocate nodes of the new component,
\PPL moves to partition where no component is split, that minimizes the distance to the initial partition $P_I$.

@@

Let $\Delta$ be the cost of \OPT.
\PPL never moves to a~partition that is more than $ \Delta$   migrations away from~the initial configuration.

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of a single repartition is at most $2\cdot\OPT$.
	The corollary is that \PPL is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{lemma}

@@

This modification alone is insufficient to obtain $O(\ell)$-competitive algorithm, and the analysis must be further improved.
In particular, pairs of nodes that did not reach the collocation threshold $\alpha$ (called external requests) incur the cost $O(\ell^2)$ for the algorithm in each phase.
The novel part of the analysis lower-bounds the cost of \OPT on external requests while considering its savings from migrations and possibly different configuration at the beginning of the phase.
This way, we show that \OPT paid a significant portion of the algorithm's cost on external requests.

\begin{theorem}
	The proposed algorithm for $k=3$ is $O(\ell)$-competitive.
	\label{thm:k=3}
\end{theorem}

>>>>

We present the algorithm \PPL that is  asymptotically optimal for the learning problem.
It utilizes the concept of components,
introduced by Avin~et~al.~\cite{repartition-disc}.
On each inter-cluster request $\{u,v\}$, 
\PPL  merges
 the two components that contain nodes $u$ and $v$.
To collocate nodes of the new component,
\PPL moves to partition where no component is split, that minimizes the distance to the initial partition.

@@

\noindent
Let $\Delta$ be the cost of \OPT.
\PPL never moves to a~partition that is more than $ \Delta$   migrations away from~the initial configuration.

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of a single repartition is at most $2\cdot\Delta$.
	The corollary is that \PPL is $O(k\cdot\ell)$-competitive.
\end{lemma}

@@

This modification alone is insufficient to obtain $O(\ell)$-competitive algorithm: pairs of nodes that did not reach the collocation threshold $\alpha$ incur the cost $O(\ell^2)$.
We carefully analyse this cost and relate it to the cost of \OPT.

\begin{theorem}
	The  algorithm for $k=3$ is $O(\ell)$-competitive.
	\label{thm:k=3}
\end{theorem}
